---
title: "Tidy Tuesday 2024-10-08"
subtitle: "National Parks Species Data"
author: "Hanna Chang"
format: 
  html:
    embed-resources: true
---

## Loading packages

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readr)
library(sf)
library(scales)
```

## Loading data

```{r, message=FALSE, warning=FALSE}
most_visited_nps_species_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-10-08/most_visited_nps_species_data.csv')
```

With a glance at the initial data, which is huge, I'd like to explore the different species that each national park has. Maybe I can also see how diverse the national parks are.

## Data wrangling by name of national park

```{r, message=FALSE, warning=FALSE}
by_parkname <- most_visited_nps_species_data %>%
  group_by(ParkName) %>%
  summarise(total_species = n_distinct(SciName)) %>%
  arrange(desc(total_species))

head(by_parkname, 10)
```

## More data wrangling for each national park's species, grouped by category name

```{r, message=FALSE, warning=FALSE}
by_category <- most_visited_nps_species_data %>%
  group_by(CategoryName) %>%
  summarise(total_species = n_distinct(SciName)) %>%
  arrange(desc(total_species))

join_category <- most_visited_nps_species_data %>%
  left_join(by_category, by = "CategoryName") %>%
  group_by(ParkName, CategoryName) %>%
  summarise(total_species = n_distinct(SciName)) %>%
  arrange(desc(total_species))

join_category <- join_category %>%
  group_by(ParkName) %>%
  mutate(percentage = total_species / sum(total_species) * 100)
```

## Each national park's species-diversity data viz

```{r, fig.width=20, fig.height=12, message=FALSE, warning=FALSE}
ggplot(join_category, aes(x = 2, y = percentage, fill = CategoryName)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y") +
  xlim(0.5, 2.5) +
  facet_wrap(~ParkName, ncol = 5, labeller = label_wrap_gen(width = 15)) +
  theme_minimal(base_size = 14) +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    panel.grid = element_blank(),
    legend.position = "right",
    legend.text = element_text(size = 10, face = "italic"),
    legend.title = element_text(size = 12, face = "bold"),
    strip.text = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold"),
    panel.spacing = unit(2, "lines"),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10),
    caption = element_text(size = 12, face = "italic"),
    plot.caption = element_text(hjust = 0, face = "italic")
  ) +
  #scale_colour_gradientn(colours = terrain.colors(16)) +
  scale_fill_viridis_d(option = "F") +
  labs(title = "Each National Park's Species Diversity by Category",
       caption = "Data Source: National Parks Species Data",)
```

Based on the visual, I'd like to see if species diversity has anything to do with the park's visitor count. I could investigate that in the future.

